---
title: "Danmap - Er interaktivt kort over steder i Danmark"
author: "Mette Pilgaard"
date: "`r Sys.Date()`"
output: html_document
---


###   GETTING STARTED WITH LEAFLET

# Try to work through down this script, observing what happens in the plotting pane.

# Review favorite backgrounds in:
# https://leaflet-extras.github.io/leaflet-providers/preview/
# beware that some need extra options specified

```{r}
# To install Leaflet package, run this command at your R prompt:
#install.packages("leaflet")

# We will also need this widget to make pretty maps:
#install.packages("htmlwidget")

# Activate the libraries
library(leaflet)
library(htmlwidgets)
```


########## Example with Markers on a map of Europe


# First, create labels for your points
```{r}
popup <-  c("Robin", "Jakub", "Jannes")
```


# You create a Leaflet map with these basic steps: you need to run the whole chain of course

```{r}
leaflet() %>%                                 # create a map widget by calling the library
  addProviderTiles("Esri.WorldPhysical") %>%  # add Esri World Physical map tiles
  addAwesomeMarkers(lng = c(-3, 23, 11),      # add layers, specified with longitude for 3 points
                    lat = c(52, 53, 49),      # and latitude for 3 points
                    popup = popup)            # specify labels, which will appear if you click on the point in the map

```


### Let's look at Sydney with setView() function in Leaflet

```{r}
leaflet() %>%
  addTiles() %>%                              # add default OpenStreetMap map tiles
  addProviderTiles("Esri.WorldImagery",       # add custom Esri World Physical map tiles
                   options = providerTileOptions(opacity=0.5)) %>%     # make the Esri tile transparent
  setView(lng = 151.005006, lat = -33.9767231, zoom = 10)              # set the location of the map 

```


```{r}
# Now let's refocus on Europe again
leaflet() %>% 
  addTiles() %>% 
  setView( lng = 2.34, lat = 48.85, zoom = 5 ) %>%  # let's use setView to navigate to our area
  addProviderTiles("Esri.WorldPhysical", group = "Physical") %>% 
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>% 
  addProviderTiles("MtbMap", group = "Geo") %>%
  addLayersControl(                                
  baseGroups = c("Geo","Aerial", "Physical"),
  options = layersControlOptions(collapsed = T))
```

 # we are adding layers control to the maps


# click the box in topright corner in your Viewer 
# to select between different background layers


########## SYDNEY HARBOUR DISPLAY WITH LAYERS
# Let's create a more complicated map 

# Set the location and zoom level

```{r}
leaflet() %>% 
  setView(151.2339084, -33.85089, zoom = 13) %>%
  addTiles()  # checking I am in the right area

```


# Bring in a choice of esri background layers  

# Create a basic basemap

```{r}
l_aus <- leaflet() %>%   # assign the base location to an object
  setView(151.2339084, -33.85089, zoom = 13)

```


# Now, prepare to select backgrounds

```{r}
esri <- grep("^Esri", providers, value = TRUE)

```


# Select backgrounds from among provider tiles. To view them the options, 
# go to https://leaflet-extras.github.io/leaflet-providers/preview/
for (provider in esri) 

```{r}
for (provider in esri) {
  l_aus <- l_aus %>% addProviderTiles(provider, group = provider)
}

```


### Map of Sydney, NSW, Australia
# We make a layered map out of the components above and write it to 
# an object called AUSmap

```{r}
AUSmap <- l_aus %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
addControl("", position = "topright")
```



# run this to see your product

```{r}
AUSmap
```



########## SAVE THE FINAL PRODUCT

# Save map as a html document (optional, replacement of pushing the export button)
# only works in root

```{r}
saveWidget(AUSmap, "AUSmap.html", selfcontained = TRUE)
```



###################################  YOUR TASK NUMBER one

# OPGAVE - Danmap - Et interaktivt kort med steder i Danmark 




# Task 1: Create a Danish equivalent of AUSmap with esri layers, 
# but call it danmap

# Lav danmap 

## Lav et basiskort med base i Odder

```{r}
leaflet() %>% 
  setView(10.0673861,55.9979681,zoom = 13) %>%
  addTiles()  # checking I am in the right area

l_dk <- leaflet() %>%   # assign the base location to an object
  setView(10.0673861,55.9979681,  zoom = 13)
```


# Now, prepare to select backgrounds

```{r}
esri <- grep("^Esri", providers, value = TRUE)
```


# Select backgrounds from among provider tiles. To view them the options, 
# go to https://leaflet -extras.github.io/leaflet-providers/preview/
for (provider in esri)

```{r}
{
 
   l_aus <- l_aus %>% addProviderTiles(provider, group = provider)
}

danmap <- l_dk %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
  addControl("", position = "topright")
```


########## ADD DATA TO LEAFLET

# In this section you will manually create machine-readable spatial
# data from GoogleMaps: 

### First, go to 
### Enter the coordinates of your favorite leisure places in Denmark 
      # extracting them from the URL in googlemaps, adding name and type of monument.
      # Remember to copy the coordinates as a string, as just two decimal numbers separated by comma. 

# Caveats: Do NOT edit the grey columns! They populate automatically!

### Second, read the sheet into R. You will need gmail login information. 
      # watch the console, it may ask you to authenticate or put in the number 
      # that corresponds to the account you wish to use.

# Libraries indlæses

```{r}
library(tidyverse)
library(googlesheets4)
library(leaflet)
```


# Read in a Google sheet

```{r}
places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
                     sheet = "DM2023",
                     col_types = "cccnncnc")
glimpse(places)
```


# load the coordinates in the map and check: are any points missing? Why?
Der er ingen points, der mangler. 

```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description)


```


#########################################################


# Task 2: Read in the googlesheet data you and your colleagues 
# populated with data into the DANmap object you created in Task 1.

```{r}
danmap <- leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description)
```

```{r}
danmap
```


# Task 3: Can you cluster the points in Leaflet? Google "clustering options in Leaflet"

```{r}
danmap <- leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description,
             clusterOptions = markerClusterOptions() )

danmap
```



# Task 4: Look at the map and consider what it is good for and what not.

## Alle steder, der skal vises på kortet, klumper, men ved at zoom ind kan man se de enkelte steder. 
Det fungerer godt med at trykke på de steder, som blev clustret i opgave 3, 
hvorved de bliver spredt ud på et udzoomet kort. 

###Men der mangler en beskrivelse af, havd der er at se på de enkelte steder. Det er blevet gjort i opgave 5, men det er en dårlig ting,
at man skal trykke på kortet for at se, hvad der er på det enkelte sted. En label, hvor det beskrives hvad det er for en sted, 
burde vises ved det enkelte punkt.



# Task 5: Find out how to display notes and classifications in the map.

# Placenames (stednavne) sættes ind på kort

```{r}
danmap <- leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description,
             clusterOptions = markerClusterOptions(), label = places$Placename )
danmap
```



## Noter - beskrivelse af sted sættes ind på kort

```{r}
danmap <- leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description,
             clusterOptions = markerClusterOptions(), label = paste(places$Placename, "-", places$Type, "-", places$Notes))

danmap



```


# se danmap

```{r}
danmap
```


# Gemme danmap

```{r}
saveWidget(danmap, "danmap.html", selfcontained = TRUE)
```




